---
title: "Reproducible Report Template - Basic"
author: "Student/s SID"
subtitle: "Project X"
date: "University of Sydney | Unit of Study | MONTH YEAR"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

<br>

# Executive Summary
- The aim of this report is to ...
- The main discoveries are ...

<br>

# Full Report

## Initial Data Analysis (IDA)

```{r}
shootings = read.csv("fatal-police-shootings-data.csv", header = TRUE)

# Quick look at top 5 rows of data
head(shootings)

## Size of data
dim(shootings)

## R's classification of data
class(shootings)

## R's classification of variables
str(shootings)
```
```{r}
shootings_copy = read.csv("fatal-police-shootings-data.csv", header = TRUE)
```

```{r}
shootings_copy$race = gsub("A", "Asian", shootings_copy$race)
shootings_copy$race = gsub("B", "Black", shootings_copy$race)
shootings_copy$race = gsub("H", "Hispanic", shootings_copy$race)
shootings_copy$race = gsub("N", "N.American", shootings_copy$race)
shootings_copy$race = gsub("W", "White", shootings_copy$race)
shootings_copy$race = gsub("O", "Other", shootings_copy$race)
shootings_copy$race[shootings_copy$race == ""] <- "N/A"
```
The data was duplicated to allow us to manipulate it and reformat it without influencing the original raw data. Variables such as body_camera and signs_of_mental_health were both reclassified into logical values as the only contained the True of False values. The name, date, manner_of_deat, armed, age, gender, race, city, state, threat_level, flee and is_geocoding_exact were all changed from characters to factors.




```{r}
shootings_copy$date = as.factor(shootings_copy$date)
shootings_copy$manner_of_death = as.factor(shootings_copy$manner_of_death)
shootings_copy$armed = as.factor(shootings_copy$armed)
shootings_copy$gender = as.factor(shootings_copy$gender)
shootings_copy$race = as.factor(shootings_copy$race)
shootings_copy$city = as.factor(shootings_copy$city)
shootings_copy$state = as.factor(shootings_copy$state)
shootings_copy$signs_of_mental_illness = as.logical(shootings_copy$signs_of_mental_illness)
shootings_copy$threat_level = as.factor(shootings_copy$threat_level)
shootings_copy$flee = as.factor(shootings_copy$flee)
shootings_copy$body_camera = as.logical(shootings_copy$body_camera)
shootings_copy$is_geocoding_exact = as.factor(shootings_copy$is_geocoding_exact)

str(shootings_copy)
```

Summary:

- The data came from ...
- The data is/is not valid because ...
- Possible issues include ...
- Each row represents ...
- Each column represents ...

<br>

## Research Question 1

```{r}
race = shootings_copy$race
table(race)

race = factor(race, levels=c("Other", "N.American", "Asian", "N/A", "Hispanic", "Black", "White"))
barplot(table(race), main="Race vs Number of Deaths", xlab="Race", ylab="Number of Deaths")
```

Insert text and analysis. 

Summary: 

<br>

## Research Question 2

Insert text and analysis. 

Summary: 

<br>

# References (if needed)

Style: APA 

<br>
